---
title: "Lab3"
author: "Sylvia Baeyens"
date: "9/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r, echo= FALSE, include= FALSE}
#including necessary libraries
library(data.table)
library(dplyr)
library(R.utils)
```

Initial question= Examine Windspeed & Temp at Station with Highest Elevation

#Reading in Data

```{r}
#download file ONLY IF it has not been downloaded previously
if (!file.exists("met_all.gz")) {
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)}
met <- data.table::fread("met_all.gz")
```

#Checking dimensions 

```{r}
dim(met)
head(met)
tail(met)
```

There are 2,377,343 rows & 30 columns in the met dataset.

# Looking at Variables & Key Variables & Data Clean-up

```{r}
str(met)

table(met$year)
table(met$day)
table(met$hour)
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)

#replacing 9999.0 elevation with N//A
met[met$elev==9999.0] <- NA
summary(met$elev)

#now the highest elevation is at 4113m

#removing temperatures of -40C
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)

#going another step further and removing temperatures less than -15C
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

#Summary Statistics

```{r}
#examining station with highest elevation
elev <- met[elev==max(elev)]
summary(elev)

#correlation between temp & wind speed
cor(elev$temp, elev$wind.sp, use="complete")
#correlation between temp & hour
cor(elev$temp, elev$hour, use="complete")
#correlation between wind speed & day
cor(elev$wind.sp, elev$day, use="complete")
#correlation between wind speed & hour
cor(elev$wind.sp, elev$hour, use="complete")
#correlation between temp & day
cor(elev$temp, elev$day, use="complete")

```

#Exploratory Graphs

```{r}
hist(met$elev, breaks=100)

hist(met$temp)

hist(met$wind.sp)

#looking for where station with highest elevation is located -> in CO Rocky Mountain Park
library(leaflet)
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)

#finding time series of temp and wind speed
#creating date-time variable
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
elev <- elev[order(date)]
head(elev)

#time series for temp
plot(elev$date, elev$temp, type='l')

#time series for wind speed
plot(elev$date, elev$wind.sp, type='l')
```

Temperature fluctuates throughout each day, with the maximum (btwn 10 & 15C) and minimum (btwn 2 & 6C) temperatures per day remaining relatively constant between days. It is likely that the max temp each day is around noon and the minimum at midnight, but would need clearer time stamps (including hours) to verify.

Wind changes much more randomly during and between days. It was very windy between August 12th and August 18th and between the 24th and 27th. 
